Imputation of incomplete multilevel data

### 2024-02-29 manuscript.qmd
`mice` arguments for multilevel data

@@

>>>>

# Methods

The [R]{.proglang} package [mice]{.pkg} provides a framework for imputing incomplete data on a variable-by-variable basis. The [mice]{.fct} function allows users to flexibly specify how many times and under what model the missing data should be imputed. This is reflected in the first four function arguments

```r
mice(data, m, method, predictorMatrix, ...)
```

where `data` refers to the incomplete dataset, `m` determines the number of imputations, `method` denotes the functional form of the imputation model per variable and `predictorMatrix` specifies the interrelational dependencies between the variables and imputation models (i.e., the set of predictors to be used for imputing each incomplete variable).

The object supplied as `data` should be tabular (e.g. a `data.frame` with $n$ rows and $p$ variables, with missing values coded as `NA`). For multilevel imputation models, a numeric clustering variable is required.

The number of imputations `m` should be determined based on the severity of the missing data problem and the intended analysis model of substantive interest. Van Buuren (2018, $\S$ 2.8) suggests using the default `m = 5` for imputation model building, and to increase `m` as required after initial exploration.

The `method` argument specifies the imputation method to be used for each column in data. If not supplied by the user, `method` defaults to convenient standard methods for single level continuous and categorical data. Since these do not take any clustering or multilevel structures into account, valid imputation of incomplete multilevel data will typically require a user-supplied methods vector. The tables 7.2, 7.3 and 7.4 in van Buuren (2018, $\S$ 7.6, $\S$ 7.7 and $\S$ 7.8, respectively) provide an overview of the available methods to perform univariate multilevel imputation.

With the `predictorMatix` argument, `mice` users can define which columns should be used as predictors in each imputation model. The default predictor matrix is a square binary matrix with the variables to be imputed in the rows and the imputation model predictors in the columns. The default `predictorMatrix` will not be suitable for multilevel data. Univariate imputation methods for two-level data use other codes than `0` and `1`. In the predictor matrix, `-2` denotes the class variable, a value `1` indicates a fixed effect and a value `2` indicates a random effect. Additionally, the method `2l.pan` uses codes `3` and `4` to add class means to codes `1` and `2` respectively.
<!--  -->
<!-- 2l.bin, 2l.lmer, 2l.norm, 2l.pan, 2lonly.mean, 2lonly.norm and 2lonly.pmm use code -2 to indicate the class variable --> -->
<!-- -	2l.bin, 2l.lmer, 2l.norm and 2l.pan use code 2 to indicate the random effects -->
<!-- -	2l.pan uses codes 3 and 4 to add class means to codes 1 and 2 respectively -->

FIMD, section 7.10

@@

# References
